Carbon Nucleosynthesis

### 2023-03-30 thesis/tex/main.tex
better yield plots

\documentclass[12pt,oneside]{book}

@@

\titleformat{\chapter}{\normalfont\normalsize\centering}{Chapter \thechapter:}{1em}{}
\titleformat{\section}{\normalfont\normalsize\centering}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\centering}{\thesubsection}{1em}{}
\titlespacing*{\chapter}{0pt}{-15pt}{40pt} \titlespacing*{\section}{0pt}{0pt}{20pt}
\renewcommand{\contentsname}{Table of Contents}

@@

\usepackage{amsmath}	% Advanced maths commands
\usepackage{txfonts}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,
    urlcolor=blue,
    pdftitle={Thesis},
    pdfpagemode=FullScreen,
}
% \usepackage{amssymb}	% Extra maths symbols

@@

\usepackage{natbib}
\usepackage[nottoc,numbib]{tocbibind}

% \usepackage{fancyhdr}
% \pagestyle{fancy}
% \fancyhead{}
% \fancyhead[RO,RE]{Carbon Galactic Nucleosynthesis}
\pagestyle{plain}

@@

\begin{titlepage}
   \begin{center}
       \vfill
       \textbf{The Galactic Chemical Evolution of Carbon} \\
        Implications for nucleosynthesis
        \vfill
        Research Thesis
        \vfill
        Presented in partial fulfillment of the requirements for graduation \textit{with research distinction} in Astronomy and Astrophysics in the College of Arts and Sciences of The Ohio State University.
        \vfill
        by 
        \vfill
       {Daniel A. Boyea}\\
       \vfill
       Department of Astronomy\\
       The Ohio State University\\
       \vfill
       April 2023
       \vfill
       Thesis Committee:\\
       David Weinberg, Advisor \\
       Jennifer Johnson \\
       Rolando Vald\'es Aguilar\\
       \vfill

@@

% Abstract of the paper
\chapter*{Abstract}
\begin{center}
Stellar evolution models provide highly uncertain predictions for elemental yields due to our limited understanding of stellar physics. In this paper, we aim to estimate the nucleosynthetic yields of carbon using galactic chemical evolution models.

We find that the AGB stars make up a fraction $0.20_{-20}^{+80}$ of total carbon production at late times. 
\end{center}

\chapter*{Acknowledgements}

Many thanks to

@@

Where does each chemical element come from?

Stellar evolution models can be used to predict how much of each element stars with a variety of physical properties produce. However, these models are rife with uncertainties, limited by our understanding of critical physical processes such as nuclear reaction rates, convection, opacity, and mass loss (CITE, e.g. stellar physics reviews).

@@

However, theoretical interpretations of the trends of

@@

\begin{figure}[htp]
    \centering
    \includegraphics[scale=0.85]{subgiants_mgh.pdf}
    \includegraphics[scale=0.85]{subgiants_mgfe.pdf}
    \caption[APOGEE Subgiants]{The [C/Mg] ratio against [Mg/H] (left) and [Mg/Fe] (right). The left panel uses blue and orange to represent high and low $\alpha$ sequences respectively. On the right, we plot more metal rich ([Mh/H]) stars with darker colors.}
    \label{fig:subgiants}
\end{figure}
\chapter{Nucleosynthesis}

Modern theoretical nucleosynthesis provides our starting point for our models.  We focus on three primary nucleosynthetic pathways: AGB stars, CCSne, and SNeIa. Each process has unique timescales and element production which we can trace through the tools of GCE. While carbon is only known to be produced in AGB stars and CCSNe, comparing carbon to iron production (from SNeIa) adds a valuable constraint to our models and insight into the delay-time production of carbon.

@@

    \caption{The net fraction carbon yield ($M_\text{C}$, the amount of carbon produced by the star divided by $M$) plotted as a function of mass, $M$. Lighter colors represent higher values of metalicity ([M/H]). Each panel represents a different AGB model. K10 reports stars at MoverH -3, -2, etc. I can reproduce C11 and K10 but not other V13 and KL16 :(((}

@@

    \caption{On the left, we plot the IMF weighted net fractional carbon yields of each AGB model as a function of metalicity.The right panel plots the normalized net carbon AGB yield as a function of time for a population of stars born at the same time at solar metallicity. We note that for the right panel, the minimum of the DTD for V13 is }

@@

\begin{figure}[htp]
    \includegraphics[scale=0.8]{y_c_cc-2.pdf}
    \includegraphics[scale=0.8]{y_co_cc.pdf}
    \caption{The fist plot shows IMF-weighted net fractional CCSNe yield of
        carbon, $y_C^{cc}$ against metalicity, [M/H], for different studies.
        The right plot instead shows [C/O]$^\text{CC}$, the logarithm of the
        abundance ratio of carbon to oxygen if CCSNe were the only carbon
        source, also as a function of metallicity for the same set of studies.
        The black line represents the carbon yield of the fiducial model,
        $y_C^{CC} = 0.004? (Z/Z_{\sun})^{0.3}$, and becomes dashed over regions
        where we have not tested our model.\citep{NKT13}, \citep{LC18},
        \citep{sukhbold+16}, \citep{WW95}
 }
    \label{fig:y_cc}
\end{figure}

While there are many stellar models providing predictions of CCSNe yields, the results of these models are highly uncertain due to the many stellar modeling uncertainties.

@@

\begin{figure*}
\label{fig:c_evo}
\includegraphics{evo_tracks.pdf}
\includegraphics{evo_slices.pdf}
\caption{The top set of panes show the time evolution of the gas phase of our
    fiducial model, plotted as tracks through \caah and \caafe where each colored line corresponds to the evolution at that galactic radius. 
The bottom set instead shows the galaxies gas-phase \caah and \caafe trend at 5 different time slices. 
}
\end{figure*}

@@

\begin{figure*}
\includegraphics[scale=0.9]{oob_agb.pdf}

\caption{Left: the \caah~ \textbf{median} present-day stellar tracks for four different
AGB models (\cristallo, \karakas, \kl, and \ventura ). The median bin [C/Mg] values from \citet{jack_subgiant} are
plotted as black points with 1-std errors marked by horizontal grey dashes.
Right: same as left but for \caafe. While each AGB yield model may
differ in detail, there is minimal effect on the overall [C/Mg] trends. }
\label{fig:agb_sims}
\end{figure*}

@@

\begin{table}
	\centering
	\caption{For each AGB yield set, our calculation of the effective SFH- and IMF-weighted carbon AGB yield, along with the multiplicative factor to reach an AGB contribution of 20.}
	\label{tab:alpha_agb}
	\begin{tabular}{lcr} % four columns, alignment for each
		\hline
		Study & $y_{\rm C, 0}^{\rm agb}$ & $\alpha_{\rm agb, 20}$\\
		\hline
		C11 & 0.000347 & 2.9\\
		K10 & 0.000585 & 1.7\\
		V13 & 0.000060 & 16.5\\
		K16 & 0.000421 & 2.4\\
		\hline
	\end{tabular}
\end{table}

@@

\begin{figure*}\label{fig:beta_f}
\includegraphics{beta.pdf}
\includegraphics{f_agb.pdf}

\caption{Similar to \ref{fig:agb_sims} except the top plot shows the fiducial model with lower and higher values of $\beta$, the C-CCSNe metallicity dependence. The bottom plot is the same except shows varying AGB fractions.}
\end{figure*}

@@

\begin{figure*}
\includegraphics[]{lateburst_eta.pdf}

\caption{Same as Fig. \ref{fig:agb_sims} but comparing the fiducial model to the reduced outlow model and a lateburst model. (okay, need to totally change approach...)}
\end{figure*}

@@

\begin{figure*}
    \includegraphics[scale=0.9]{cooh_scatter.pdf}
    \includegraphics[scale=0.9]{cnoh_scatter.pdf}
    \caption{The simulated stars of our model at present day plotted on \caah~and colored such that lighter colors represent stars born at greater galactric radii.}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[]{summary-2.pdf}
\caption{Present-day gas phase tracks of the fiducial model in \caah~space. M101 is valid comparison to MW ...(van Dokkum et al. 2014)}
\label{fig:gas_phase}
\end{figure*}

@@

\begin{figure}
    \centering
    \includegraphics[]{analytic.pdf}
    \caption{Our reverse fitting method to determine total CCSNe contribution}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Data Availability}

@@

\bibliographystyle{aasjournal}
\bibliography{main} % if your bibtex file is called example.bib

@@

\appendix

\chapter{Validating the APOGEE Subgiant Sample}\label{sec:jack}
Our APOGEE subgiant sample excludes stars marked with the following flags:
\begin{itemize}
\item \texttt{ancillary young embedded cluster}
\item \texttt{ancillary emission line star}
\item \texttt{MIR-deected candidate cluster member}
\item \texttt{selected as part of the EB program}
\item \texttt{selected as part of the young cluster study (in-SYNC)}
\item \texttt{W3/4/5 star forming complex}
\end{itemize}

@@

\chapter{Effects of Alternate AGB Models}\label{sec:alt_agb}

@@

\begin{figure*}
    \includegraphics{coofe_agb_extra.pdf}
    \includegraphics{coofe_agb_extra_0.2.pdf}

\caption{Same as \ref{fig:agb_sims} but comparing our fiducial model to our recommended z-dependent $y_\text{C}^\text{CC}$ model. A plot of the stars of our favored CCSNe model colored by radius. The contours represent the density of V+21 data. The model is given Gaussian scatter equivalent to the mean error in the abundances reported by APOGEE. }
\end{figure*}

>>>>

\documentclass[12pt,oneside]{report}

@@

\titleformat{\chapter}{\normalfont\normalsize\centering}{Chapter \thechapter:}{1em}{}
\titleformat{\section}{\normalfont\normalsize\centering}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\centering}{\thesubsection}{1em}{}
\titlespacing*{\chapter}{0pt}{0pt}{40pt} \titlespacing*{\section}{0pt}{0pt}{20pt}
\renewcommand{\contentsname}{Table of Contents}
\AtBeginDocument{
      \renewcommand{\bibsection}{\chapter*{\bibname}}
  }

@@

\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
\makeatletter
% \patchcmd{<cmd>}{<search>}{<replace>}{<succes>}{<failure>}
\patchcmd{\@chapter}{\addtocontents{lof}{\protect\addvspace{10\p@}}}{}{}{}% LoF
\patchcmd{\@chapter}{\addtocontents{lot}{\protect\addvspace{10\p@}}}{}{}{}% LoT
\makeatother

\usepackage{amsmath}	% Advanced maths commands
\usepackage{txfonts}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,
    urlcolor=blue,
    citecolor=blue,
    pdftitle={Thesis},
    pdfpagemode=FullScreen,
}
% \usepackage{amssymb}	% Extra maths symbols

@@

\usepackage{natbib}
\usepackage[nottoc,numbib]{tocbibind}
\usepackage[titles]{tocloft} % fig titles
\renewcommand{\cftfigpresnum}{\figurename\enspace}
\setlength{\cftfignumwidth}{5em}
\renewcommand{\cfttabpresnum}{\tablename\enspace}
\setlength{\cfttabnumwidth}{5em}

% \usepackage{fancyhdr}
% \pagestyle{fancy}
% \fancyhead{}
% \fancyhead[RO,RE]{Carbon Galactic Nucleosynthesis}
\pagestyle{plain}
\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}

@@

\begin{titlepage}
   \begin{center}
       \vfill
       \textbf{The Galactic Chemical Evolution of Carbon} \\
        Implications for nucleosynthesis
        \vfill
        Research Thesis
        \vfill
        Presented in partial fulfillment of the requirements for graduation \textit{with research distinction} in Astronomy and Astrophysics in the College of Arts and Sciences of The Ohio State University.
        \vfill
        by 
        \vfill
       {Daniel Alexandra Boyea}\\
       \vfill
       Department of Astronomy\\
       The Ohio State University\\
       \vfill
       April 2023
       \vfill
       Thesis Committee:\\
       David Weinberg, Advisor \\
       Jennifer Johnson \\
       Rolando Vald\'es Aguilar\\
       \vfill

@@

% Abstract of the paper
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
Stellar evolution models provide highly uncertain predictions for elemental yields due to our limited understanding of stellar physics. In this paper, we aim to estimate the nucleosynthetic yields of carbon using galactic chemical evolution models.

We find that the AGB stars make up a fraction $0.20_{-20}^{+80}$ of total carbon production at late times.

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

James Johnson, for guiding me through this project and through my last year of undergraduate.
Wayne Schlingman, for everything you do for everyone. You are the glue holding together our department.
David Weinberg and Jennifer Johnson, for listening to my earlier drafts of this work and providing excellent suggestions.

I also would not have made it here without all the additional support from my friends and family. Eric, even across the country, you have always been a point of support and stability, and I am always happy to be around you. Anya, you are a wonderful friend and I cannot thank you enough for your support this year. 
I love so many people in our department who have just been friends and positive support: Kaia, Maria, Autumn, Simon, Mary, Bailee, Aaliyah, Sanskruti, Alyssa, Harrison, Denis, and so many others. 
My Family (and Arya, my lovely labrodoodle).

@@

Astronomy is unique among sciences in that we can determine the history of chemical elements. While stable on earth, the churning gasses and firery stars can dramatically transform elements into new materials. However, we still have much to learn about how each chemical element is made. In this work, we will hope to clarify how the chemical element carbon, the basis of life, is made in the universe.

As the universe just began to cool following the big bang, the only elements around were hydrogen, helium, and trace amounts of lithium and berylium. All other elements were formed afterwords with the help of stars. Stars of different masses, ages, and metalicities have substantially different interior stuctures. Different stars produce vastly different amounts of different chemical elements depending on their properties.

Stellar evolution, at the basis of galactical chemical evolution, predict the elemental production of different stellar processes. 
However, these models are rife with uncertainties, limited by our understanding of critical physical processes such as nuclear reaction rates, convection, opacity, and mass loss (CITE, e.g. stellar physics reviews).

@@

However, our theoretical understanding of the producion sources of carbon are still uncertain.

@@

\begin{figure}[htp]
    \includegraphics{subgiants_mgh.pdf}
    \includegraphics{subgiants_mgfe.pdf}
    \caption[APOGEE Subgiants]{The [C/Mg] ratio against [Mg/H] (left) and [Mg/Fe] (right) for the sample of subgiants as in \cite{jack_subgiant}. The left panel uses blue and orange to represent high and low $\alpha$ sequences respectively. On the right, we plot more metal rich ([Mh/H]) stars with darker colors.}
    \label{fig:subgiants}
\end{figure}
\chapter{Nucleosynthesis}

Nucleosynthetic predictions 
Modern theoretical nucleosynthesis provides our starting point for our models.  We focus on three primary nucleosynthetic pathways: AGB stars, CCSne, and SNeIa. Each process has unique timescales and element production which we can trace through the tools of GCE. While carbon is only known to be produced in AGB stars and CCSNe, comparing carbon to iron production (from SNeIa) adds a valuable constraint to our models and insight into the delay-time production of carbon.

@@

        \caption[AGB carbon yields]{The net fraction carbon yield ($M_\text{C}$, the amount of carbon produced by the star divided by $M$) plotted as a function of mass, $M$. Lighter colors represent higher values of metalicity ([M/H]). Each panel represents a different AGB model. K10 reports stars at MoverH -3, -2, etc. I can reproduce C11 and K10 but not other V13 and KL16 :(((}

@@

    \caption[SSP AGB carbon]{On the left, we plot the IMF weighted net fractional carbon yields of each AGB model as a function of metalicity.The right panel plots the normalized net carbon AGB yield as a function of time for a population of stars born at the same time at solar metallicity. We note that for the right panel, the minimum of the DTD for V13 is }

@@

\begin{figure}[htp]
    \centering
    \includegraphics{y_c_cc.pdf}
    \caption[CCSNe carbon yields]{The fist plot shows IMF-weighted net fractional CCSNe yield of
        carbon, $y_C^{cc}$ against metalicity, [M/H], for different studies.
        The right plot instead shows [C/O]$^\text{CC}$, the logarithm of the
        abundance ratio of carbon to oxygen if CCSNe were the only carbon
        source, also as a function of metallicity for the same set of studies.
        The black line represents the carbon yield of the fiducial model,
        $y_C^{CC} = 0.004? (Z/Z_{\sun})^{0.3}$, and becomes dashed over regions
        where we have not tested our model.\citep{NKT13}, \citep{LC18},
        \citep{sukhbold+16}, \citep{WW95}
 }
    \label{fig:y_cc}
\end{figure}

Massive stars form $^{12}$C through the triple--$\alpha$ process. However, the carbon has to get out of the star to be useful. 
While there are many stellar models providing predictions of CCSNe yields, the results of these models are highly uncertain due to the many stellar modeling uncertainties.

@@

\begin{figure*}
\label{fig:c_evo}
\includegraphics{evo_tracks.pdf}
\includegraphics{evo_slices.pdf}
\caption[Carbon chemical evolution tracks]{The top set of panes show the time evolution of the gas phase of our
    fiducial model, plotted as tracks through \caah and \caafe where each colored line corresponds to the evolution at that galactic radius. 
The bottom set instead shows the galaxies gas-phase \caah and \caafe trend at 5 different time slices. 
}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics{oob_agb.pdf}

\caption[AGB GCE Models]{Left: the \caah~ \textbf{median} present-day stellar tracks for four different
AGB models (\cristallo, \karakas, \kl, and \ventura ). The median bin [C/Mg] values from \citet{jack_subgiant} are
plotted as black points with 1-std errors marked by horizontal grey dashes.
Right: same as left but for \caafe. While each AGB yield model may
differ in detail, there is minimal effect on the overall [C/Mg] trends. }
\label{fig:agb_sims}
\end{figure}

@@

\begin{table}
	\centering
    \caption[AGB net solar IMF yields]{For each AGB yield set, our calculation of the effective SFH- and IMF-weighted carbon AGB yield, along with the multiplicative factor to reach an AGB contribution of 20.}
	\label{tab:alpha_agb}
	\begin{tabular}{lcr} % four columns, alignment for each
		\hline
		Study & $y_{\rm C, 0}^{\rm agb}$ & $\alpha_{\rm agb, 20}$\\
		\hline
		C11 & 0.000347 & 2.9\\
		K10 & 0.000585 & 1.7\\
		V13 & 0.000060 & 16.5\\
		K16 & 0.000421 & 2.4\\
		\hline
	\end{tabular}
\end{table}

@@

\begin{figure}\label{fig:beta_f}
\centering
\includegraphics{beta.pdf}
\includegraphics{f_agb.pdf}

\caption[Adjusted yield models]{Similar to \ref{fig:agb_sims} except the top plot shows the fiducial model with lower and higher values of $\beta$, the C-CCSNe metallicity dependence. The bottom plot is the same except shows varying AGB fractions.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[]{lateburst_eta.pdf}

\caption[Lateburst and reduced-outflow models]{Same as Fig. \ref{fig:agb_sims} but comparing the fiducial model to the reduced outlow model and a lateburst model. (okay, need to totally change approach...)}
\end{figure}

@@

\begin{figure*}
    \includegraphics[scale=0.9]{cooh_scatter.pdf}
    \includegraphics[scale=0.9]{cnoh_scatter.pdf}
    \caption[Scatter agreement]{The simulated stars of our model at present day plotted on \caah~and colored such that lighter colors represent stars born at greater galactric radii.}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[]{summary-2.pdf}
\caption[Gas phase abundances]{Present-day gas phase tracks of the fiducial model in \caah~space. M101 is valid comparison to MW ...(van Dokkum et al. 2014)}
\label{fig:gas_phase}
\end{figure*}

@@

\begin{figure}
    \centering
    \includegraphics[]{analytic.pdf}
    \caption[Reverse fit yields]{Our reverse fitting method to determine total CCSNe contribution}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Data Availability}
\addcontentsline{toc}{chapter}{Data Availability}

@@

\newpage
\bibliographystyle{aasjournal}
\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{main} % if your bibtex file is called example.bib

@@

\appendix
\chapter*{Appendix}
\addcontentsline{toc}{chapter}{Appendix}
\renewcommand{\thesection}{A.\arabic{section}}
\renewcommand\thefigure{A.\arabic{figure}}    
\setcounter{figure}{0}

\section{Validating the APOGEE Subgiant Sample}\label{sec:jack}
Our APOGEE subgiant sample excludes stars marked with the following flags:
\begin{itemize}
\item \texttt{ancillary young embedded cluster}
\item \texttt{ancillary emission line star}
\item \texttt{MIR-deected candidate cluster member}
\item \texttt{selected as part of the EB program}
\item \texttt{selected as part of the young cluster study (in-SYNC)}
\item \texttt{W3/4/5 star forming complex}
\end{itemize}

@@

\newpage
\section{Effects of Alternate AGB Models}\label{sec:alt_agb}

@@

\begin{figure}[htp]
    \includegraphics{coofe_agb_extra.pdf}
    \includegraphics{coofe_agb_extra_0.2.pdf}

    \caption[Alternate AGB models]{Same as \ref{fig:agb_sims} but comparing our fiducial model to our recommended z-dependent $y_\text{C}^\text{CC}$ model. A plot of the stars of our favored CCSNe model colored by radius. The contours represent the density of V+21 data. The model is given Gaussian scatter equivalent to the mean error in the abundances reported by APOGEE. }
\end{figure}
